Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-26 VMCAI-revised-version/final.tex
up

The undecidability of which leads to the following proposition.

@@

\begin{proof}
\nocite{araki1976PN,dufourd1998reset}
{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to {\sc Structural existential halting} of reset-VASS,
since a run terminates w.l.o.g. if it reaches a terminal state of the form $q_f({\bf 0})$. \qed \end{proof}

@@

We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
%
Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the 
structural termination of reset-VASS.
%\mathieu{GOTO}

@@

\section{Lemma technical proofs}

@@

\section{More details on VASS}\label{appendix vass}

@@

From $k_0$ onwards,
$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
is stationnary. 
%\mathieu{
We check the finiteness of the set by comparing the finite basis of $\Bad$ and
$(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
%}
If the set is finite then we stop the procedure once we have checked for every element
whether it is possible to reach $\Safe$ from it.
If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
hold
for $k_m = \max(k_0, k_{\pi})$ with 
$k_{\pi}$ the maximum of the length of the paths from 
$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
If it is infinite then
for every $k \in \N$,
$k$-resilience does not hold, 
and we eventually find a witness that $k$-resilience does not hold.
Indeed, 
if
$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
is
infinite
then
there exists an element of the form $q(\textbf{u} )$
with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
and hence from which it is not possible at all to reach
$\Safe$ in $k$ or less steps.
\qed
\end{proof}

>>>>

{\sc Structural existential halting} is undecidable which leads to the following proposition.

@@

\begin{proof}
\nocite{araki1976PN,dufourd1998reset}
{\sc Resilience} when $\Safe= \{ q_f({\bf 0}) \}$ for some state $q_f$ is equivalent to {\sc Structural existential halting} of reset-VASS,
since a run terminates w.l.o.g. iff it reaches a terminal state of the form $q_f({\bf 0})$. \qed \end{proof}

@@

We now consider the case $\Safe = \mathop{\downarrow} \Safe$.
%
Unfortunately {\sc State-resilience} is undecidable in this case. As for {\sc Resilience} in WSTS when $\Safe = \mathop{\downarrow} \Safe$, this stems from undeciability of the 
{\sc Structural existential halting} of reset-VASS.
%\mathieu{GOTO}

@@

\section{Proofs of technical lemmas}

@@

\section{Decision procedure for VASS when $\Safe = \mathop{\downarrow} \Safe$}\label{appendix vass}

@@

From $k_0$ onwards,
$\Bad \setminus (\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))= 
\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
is stationnary. 
%\mathieu{
We check the finiteness of the set by comparing the elements from the finite basis of $\Bad$ and
$(\pred^{\leq k}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$.
%}
If the set is finite then we stop the procedure once we have checked for every element
whether it is possible to reach $\Safe$ from it.
If it is possible to reach $\Safe$ from every element then {\sc $k_m$-resilience}
hold
for $k_m = \max(k_0, k_{\pi})$ with 
$k_{\pi}$ the maximum of the length of the paths from 
$\Bad \setminus (\pred^{\leq k_0}(\mathop{\uparrow} q_{j_1}(\textbf{0}))\cup \ldots \cup \pred^{\leq k_0}(\mathop{\uparrow} q_{j_\ell}(\textbf{0}))$ to $\Safe$.
If it is infinite then
for every $k \in \N$,
$k$-resilience does not hold, 
and we eventually find a witness that $k$-resilience does not hold.
Indeed, 
if
$\Bad \setminus (\pred^{*}(\mathop{\uparrow} q_{j_1}(\textbf{0})) \cup \ldots \cup \pred^{*}(\mathop{\uparrow} q_{j_\ell}(\textbf{0})) )$
is
infinite
then
there exists an element of the form $q(\textbf{u} )$
with $\textbf{u} > \textbf{v}_{\Bad} + (k+1) \cdot (c_{\max}, \ldots, c_{\max})$ 
that is in $\Bad$ but not in any of the $\pred^{\leq k}(\mathop{\uparrow} q_{j_i}(\textbf{0}))$,
and hence from which it is not possible at all to reach
$\Safe$ in $k$ or less steps.
\qed
\end{proof}
